---
title: "Shotgun metagenomics report"
author:
- "Mattias  de Hollander (m.dehollander@nioo.knaw.nl)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
    rmd: "report.Rmd"
output: html_notebook
link-citations: true
---
# Shotgun metagenomics report for `r snakemake@config$project`


<div class="alert alert-dismissible alert-info">
<h4 class="fa fa-info-circle alert-heading">Info</h4>
<p class="mb-0">This is a document for reproducible science. The R Notebook file can be downloaded from the top left (code->download rmd)</p>
</div>


# Load Data
<div class="alert alert-dismissible alert-success">
  <p>To run this locally download the <a id="rdata" class="alert-link" href="link">RData file</a> and run the Rmd in the same directory</p>
</div>


```{r}
load(snakemake@input$rdata)
```

# Assembly results {.tabset .tabset-sticky}

## Assembly statistics {.active .tabset .tabset-sticky}

```{r}
quast
```

## Mapping
```{r}
flagstat
```

<script>
function download(text, id, name, type) {
  var file = new Blob([text], {type: type})
  var a = $(id)
  a[0].href = URL.createObjectURL(file)
  a[0].download = name
}
function dataURItoBlob(dataURI, callback) {
    // convert base64 to raw binary data held in a string
    // doesn't handle URLEncoded DataURIs - see SO answer #6850276 for code that does this
    var byteString = atob(dataURI.split(',')[1]);

    // separate out the mime component
    var mimeString = dataURI.split(',')[0].split(':')[1].split(';')[0]

    // write the bytes of the string to an ArrayBuffer
    var ab = new ArrayBuffer(byteString.length);
    var ia = new Uint8Array(ab);
    for (var i = 0; i < byteString.length; i++) {
        ia[i] = byteString.charCodeAt(i);
    }

    // write the ArrayBuffer to a blob, and you're done
    var bb = new Blob([ab]);
    return bb;
}
</script>
